\chap{Acknowledgement}

Completing this dissertation has been a journey I could never have undertaken alone. To the many individuals who supported, inspired, and uplifted me along the way, I offer my deepest gratitude.

First and foremost, I thank my parents, Yilong Wang and Xilin Wang, whose unwavering love, sacrifices, and steadfast belief in education laid the foundation for every achievement in my life. You taught me resilience, curiosity, and the value of hard work, even when the path seemed uncertain. This work is as much yours as it is mine.

To my incredible wife, Yinuo Li, thank you for being my rock and tolerating my eccentricities. You celebrated every small victory and supported me during moments of doubt. I am endlessly grateful for your partnership, faith and love.

I extend my profound appreciation to my academic advisor, Iulian Iordachita, whose wisdom, feedback, and trust in my vision guided this project from its earliest stages. To my committee members, Russell Taylor and Axel Krieger, thank you for your rigorous insights and for challenging me to think more critically. To my decade-long mentors since college, Kiju Lee and Mary Assad, thank you for taking me under your wings in moments of pain and hardship.

This work was also shaped by the camaraderie and support of colleagues and friends. Lidia Al-Zogbi, Guanyun Liu, Yangsheng Xu, Jiarong Kang, Chang Chang, and Junling Mei -- your collaboration and willingness to listen were invaluable. Special thanks to the Multi-Scale Medical Robotics Center of Hong Kong for funding my research, and the Laboratory for Computational Sensing + Robotics at Johns Hopkins University for fostering a community of curiosity and mutual growth.

To my dear friends, Yufan Zhang, Yuan Gao, and Zi'Ang Jia, thank you for the laughter, distractions, and reminders that there is life beyond the dissertation. You kept me sane and reminded me of the joy in the journey.

Finally, to the countless mentors, participants, and institutions who contributed indirectly to this work -- I acknowledge your role in shaping my academic path. While I cannot name you all, please know your impact is deeply felt.
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:
